---
title: "PA1_template"
author: "Hazem Haffouz"
date: "2023-12-25"
output: pdf_document
---

# Load required libraries
```{r library, echo=TRUE}
library(ggplot2)
library(dplyr)
```

# Load data
```{r cars, echo=TRUE}
activity_data <- read.csv("activity.csv")
```

# Calculate the total number of steps taken per day
```{r pressure, echo=TRUE}
total_steps_per_day <- activity_data %>%
  group_by(date) %>%
  summarise(total_steps = sum(steps, na.rm = TRUE))
```


# Histogram of the total number of steps taken each day
```{r Histogram 1, echo=TRUE}
ggplot(total_steps_per_day, aes(x = total_steps)) +
  geom_histogram(binwidth = 1000, fill = "blue", color = "black") +
  labs(title = "Total Number of Steps Taken Each Day",
       x = "Total Steps",
       y = "Frequency")
```


# Calculate and report the mean and median of the total number of steps taken per day
```{r mean medina, echo=TRUE}
mean_steps <- mean(total_steps_per_day$total_steps)
median_steps <- median(total_steps_per_day$total_steps)
```


# Average Daily Activity Pattern
```{r average, echo=TRUE}
average_daily_pattern <- activity_data %>%
  group_by(interval) %>%
  summarise(average_steps = mean(steps, na.rm = TRUE))
```


# Time series plot of the average number of steps taken
```{r Time series, echo=TRUE}
ggplot(average_daily_pattern, aes(x = interval, y = average_steps)) +
  geom_line(color = "red") +
  labs(title = "Average Number of Steps Taken in Each 5-Minute Interval",
       x = "Interval",
       y = "Average Number of Steps")
```


# Identify the 5-minute interval that has the maximum number of steps on average
```{r max interval, echo=TRUE}
max_interval <- average_daily_pattern[which.max(average_daily_pattern$average_steps), ]$interval
```


# Imputing missing values
# Strategy: Replace NA with mean for that 5-minute interval
``` {r missing replace, echo=TRUE}
imputed_data <- activity_data
for(i in 1:nrow(imputed_data)) {
  if(is.na(imputed_data$steps[i])) {
    imputed_data$steps[i] <- average_daily_pattern[average_daily_pattern$interval == imputed_data$interval[i], ]$average_steps
  }
}
```


# Histogram of the total number of steps taken each day after imputing missing values
``` {r Total number of steps, echo=TRUE}
total_steps_per_day_imputed <- imputed_data %>%
  group_by(date) %>%
  summarise(total_steps = sum(steps))
```

``` {r Histogram, echo=TRUE}
ggplot(total_steps_per_day_imputed, aes(x = total_steps)) +
  geom_histogram(binwidth = 1000, fill = "green", color = "black") +
  labs(title = "Total Number of Steps Taken Each Day (Imputed Data)",
       x = "Total Steps",
       y = "Frequency")
```


# Are there differences in activity patterns between weekdays and weekends?
# Create a new factor variable
``` {r Differences in activity, echo=TRUE}
imputed_data$date <- as.Date(imputed_data$date)
imputed_data$day_type <- ifelse(weekdays(imputed_data$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
```


# Average steps taken per interval across weekdays and weekends
``` {r Average steps, echo=TRUE}
average_steps_by_day_type <- imputed_data %>%
  group_by(interval, day_type) %>%
  summarise(average_steps = mean(steps))
```

# Panel plot
``` {r Panel, echo=TRUE}
ggplot(average_steps_by_day_type, aes(x = interval, y = average_steps, color = day_type)) +
  geom_line() +
  facet_wrap(~ day_type, ncol = 1, scales = "free_y") +
  labs(title = "Average Number of Steps Taken in Each 5-Minute Interval by Day Type",
       x = "Interval",
       y = "Average Number of Steps",
       color = "Day Type")
```